---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
head(data)
```



## What is mean total number of steps taken per day?

```{r}
library(ggplot2)
total_steps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
qplot(total_steps, xlab = "Total steps taken per day")
mean(total_steps, na.rm = TRUE)
median(total_steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

```{r}
#1 Time series plot.
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps),by=list(interval=data$interval),FUN=mean,na.rm=TRUE)
ggplot(data=averages, aes(interval,steps)) + geom_line() + xlab("5 minute interval") + ylab("Average steps taken")
#2 5 minute interval with maximum steps.
averages[which.max(averages$steps),]
```

## Imputing missing values

```{r}
#1 Total missing values.
missing <- is.na(data$steps)
table(missing)
#2 Filling missing values with mean of the respective 5 minute interval. 
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
#3 New dataset with fillled missing values.
filled.data <- data
filled.data$steps <- mapply(fill.value, filled.data$steps, filled.data$interval)
#4 New histogram, mean and median.
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```
The mean and median values obtained from the new imputed data are higher than those calculated above in the first task. This is because the missing 'NA' values have been imputed. Hence, the mean and median values now obtained are based on calculations that use mean of steps in 5 minute interval instead 'NA' values.


## Are there differences in activity patterns between weekdays and weekends?
```{r}
#1 Function to determine the type of day.
typeDay <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN=typeDay)
#2 Plotting average steps taken on weekdays and weekends.
averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5 minute interval") + ylab("Number of steps")
```
